---
title: "Reproduceable data course project"
author: "mihanentalpo@yandex.ru"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Configure data url and current working dir:

```{r}

# data url:
url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

# Setting current work directory to the path of current script
scriptPath <- function() {
    getSrcDirectory(scriptPath);
}
try(setwd(scriptPath()))

```

## Source data

Raw input data is taken from the url: `r url`

The data contains information about number of steps walked by anonymous individuals, 
per five-minutes intervals, recorded for two months.

Load data from file, it needed, download it and unzip first:

```{r}
# If the file isn't downloaded yet, lets download it:
filename = paste(getwd(), "ActivityMonitoringData.zip", sep="/")
if (!file.exists(filename))
{
    download.file(url, destfile=filename, method="curl")
}

# If the file downloaded, but not extracted, lets extract it:
datafile = paste(getwd(), "activity.csv", "sep"="/")
if (!file.exists(datafile))
{
    unzip(filename)
}

# Read data, convert 'date' column to Date, remove NA
if (!exists("raw_data"))
{
    raw_data = read.csv(datafile, header=TRUE, sep=",")
    raw_data[, 'date'] = as.Date(raw_data[, 'date'], "%Y-%m-%d")
    n_raw_data = raw_data[!is.na(raw_data), ]
}
```

### Source data summary:

```{r}
    summary(raw_data)
```

Computing total steps per day, and mean of it:

```{r}
# Aggregate
steps_by_date = aggregate(n_raw_data$steps, by=list(date=n_raw_data$date), FUN=sum)
# Remove NA values
steps_by_date = steps_by_date[!is.na(steps_by_date$x), ]
# Set appropriate column names
names(steps_by_date) = c("date", "steps")
# Compute mean total:
mean_total = as.integer(mean(steps_by_date$steps))
# Compute median of total steps per day:
median_total = as.integer(median(steps_by_date$steps))
```


```{r}
hist(steps_by_date$steps, breaks=20, xlab="Steps per day", main="Histogram of total steps number per day")
```

### Mean and median of total numbers of steps per day:
```{r}
mean_total
median_total
```

Compute average steps per five-minute intervals:

```{r}
avg_steps_by_interval = aggregate(n_raw_data$steps, by=list(date=n_raw_data$interval), FUN=mean)
# Set appropriate column names
names(avg_steps_by_interval) = c("interval", "steps")
```

Distribution of steps per time intervals:
```{r}
plot(
    avg_steps_by_interval$interval, 
    avg_steps_by_interval$steps, type="l", 
    xlab="5-minute interval number", ylab="Average steps", 
    main="Time series of average steps per 5-minute interval"
)
```

5-minute interval containing maximum average number of steps:
```{r}
    max_interval = avg_steps_by_interval$interval[ 
        max(avg_steps_by_interval$steps) == avg_steps_by_interval$steps 
    ]

    max_interval
```

Imputing NA values:

```{r}
    na_indexes = which(is.na(raw_data$steps))
    for (i in 1:length(na_indexes))
    {
        # replacing NA values by average of the same interval.
        interval = raw_data$interval[na_indexes[i]]
        avg_by_interval = which(avg_steps_by_interval$interval == interval)
        raw_data$steps[na_indexes[i]] = avg_by_interval
    }
```

Computing total steps per day after imputing NA values:
```{r}
# Aggregate
steps_by_date2 = aggregate(raw_data$steps, by=list(date=raw_data$date), FUN=sum)
```

```{r}
hist(steps_by_date2$x, breaks=20, xlab="Steps per day", main="Histogram of total steps number per day, after imputing NA values")
```

Computing average steps per interval on weekday and weekend separatedly:
```{r}
# Computing boolean vector with flags of weekday (not weekend)
wd = as.POSIXlt(n_raw_data$date)$wday + 1 < 6
avg_steps_by_interval_weekday = aggregate(
    n_raw_data$steps[wd], by=list(interval=n_raw_data$interval[wd]), FUN=mean
)
avg_steps_by_interval_weekend = aggregate(
    n_raw_data$steps[!wd], by=list(interval=n_raw_data$interval[!wd]), FUN=mean
)
```

```{r}

plot(
    avg_steps_by_interval_weekday$interval, 
    avg_steps_by_interval_weekday$x, type="l", 
    xlab="5-minute interval number", ylab="Average steps", 
    main="Time series of average steps per 5-minute interval on WEEKDAYS",
    col="red"
)
legend(1, 95, legend=c("Weekdays", "Weekends"),
       col=c("red", "blue"), lty=c(1,1), cex=0.8)

lines(
    avg_steps_by_interval_weekend$interval, 
    avg_steps_by_interval_weekend$x, type="l", 
    xlab="5-minute interval number", ylab="Average steps", 
    main="Time series of average steps per 5-minute interval on WEEKENDS",
    col="blue"
)
